Probabilistic Topic Modelling for Spatial Smoothing in Mass Spectrometry Imaging

### 2016-12-17 docs/MSci_proposal/mprop.tex
Finishing the evaluation section.

% Intro
% - ML in biology
% - MSI
% - Spatial Smoothing
\par This proposal is focused on applying machine learning techniques to improve the analysis of biology data sets. We intend to enhance the quality of \textit{mass spectrometry imaging} (MSI) data. Briefly, by the use of MSI, the researchers are able to interpret the formation of chemical processes. However, as suggested by a recent overview by Smith et al. \cite{smith_2004}, the pre-processing of molecular-level entity data sets is lagging in terms of the current noise reduction algorithm effectiveness. We will attempt to tackle this issue of noise by utilising spatial smoothing.

@@

% Objectives
% - Research in the state-of-the-art methodology.
% - Reducing the impact of noise in metabolomics.
% - Providing a general-purpose machine learning model.
\par The primary goal of this research is to expand the consensus on whether topic modelling is a valid approach to apply spatial smoothing in MSI data sets. To be more specific, we set our objectives to the following:
\begin{enumerate}
    \item Optimise the general topic modelling methods for MSI data sets;
    \item Successfully apply spatial smoothing in MSI data sets;
    \item Give a basis to a general topic modelling method for error correction.
\end{enumerate}
Briefly, the first objective is expected to utilise the general state-of-the-art topic modelling methodology to distinguish the techniques displaying better performance in MSI applications -- we will conduct a survey on LDA-like topic modelling methodology. The second objective focuses on optimising those general models for the particular task of spatial smoothing in MSI data sets. Finally, the third objective can be treated as a general contribution to the field of machine learning, since the discovered spatial smoothing techniques might be applicable to other domains.

@@

% General intuition on topic modelling
% - Example
\par Recall that in topic modelling induces a hidden structure within the studied data sets. We will provide an example providing an intuition in this process. Say, if we were analysing a book, its chapters might address different topics or address the topics at a different extent. Further, the chapters itself would have unique distributions of the topics. Note that we are assuming that we do not know what the topics addressed in the book. In order to derive these hidden topics, a probabilistic topic modelling method would analyse the structure of the words used in the book. In other words, the model would induce patterns which would refer to a particular topic. Regarding the method's performance, it would depend on the method's ability to optimise the inference of the patterns. For example, commonly used words (such as and, or, the, etc.) would be relevant in all topics. Therefore, their semantic impact in defining the topics would be negligible. For this reason, we could dismiss these to improve the time performance of the method.

% LDA intro
% - the model
% - A notification that the terminology and the notation will be useful throughout the proposal.
\par The terminology used in this proposal is equivalent to the terminology used by Blei \cite{blei_2012}. Figure \ref{fig:lda} below represents the original LDA model in plate notation. We will use it as a guide to familiarise with the terminology.   
\begin{figure}[H]
  \centering
  \includegraphics[width=0.7\textwidth]{lda_model}
  \caption{The design of the original LDA model}
  \label{fig:lda}
\end{figure}

% High level LDA terminology
% - Notation:
% -- List LDA definitions
% -- Explain the definitions
\par To start with, we will familiarise with the higher level entities of the model. The circles represent an entity, whereas the plates indicate the number of entities. For example, there would $N \times M$ entities of $z$ and $M$ entities of $\theta$. Further, LDA is a three-level hierarchical structure: the smallest entity (grey circle) is defined as a \textit{word}; then a collection of words (inner plate) is defined as a \textit{document}; and a collection of documents (outer plate) is a \textit{corpus}. Note that Figure \ref{fig:lda} represents $M$ documents and $N$ words per document, it follows that in the corpus there would be $M \times N$ words. Also, as suggested by Figure \ref{fig:lda}, let's denote a word by $w$. Then, let's denote the sequence of words in document $d$ by 
\begin{equation*}
\mathbf{w}_d = \{w_1, w_2, \dots, w_N\},
\end{equation*}
then the sequence of words in the corpus would be denoted by
\begin{equation*}
\mathcal{\mathbf{D}} = \{\mathbf{w_1}, \mathbf{w_2}, \dots, \mathbf{w_M}\}.
\end{equation*}

% Low level LDA terminology
% - Intro to observable and hidden variables
% - Notation:
% -- List the definitions
% -- Explain the definitions
\par Now, we will introduce the lower level entities. The grey circle indicates that the \textit{observable} variables, whereas the white coloured circles indicate the \textit{hidden} variables. Note that in Figure 1 the words are the only observable variables; all other entities are hidden, i.e., they describe the underlying structure of the model. Further, recall that topic modelling is applied in order to discover a mixture of topics; or in other words, topic modelling could describe a corpus and its documents in terms of topic distributions. This intuition will simplify the definitions of the remaining entities in Figure 1, these are given as follows:
\begin{itemize}
\item $K$ is the number of topics;
\item $V$ is the size of the vocabulary, i.e. the number of unique words in the corpus;
\item $\alpha$ is the parameter referring to the prior distribution of the topics over documents;
\item $\beta$ is the parameter referring to the prior distribution of the vocabulary over topics;
\item $\theta$ is the latent variable referring to the topic distribution over documents;
\item $\phi$ is the latent variable referring to the vocabulary over topics
\item $w$ is the observable variable referring to the words in documents;
\item $z$ is the latent variable of topic assignments over the words.
\end{itemize}
Further, note that the listed entities can be perceived as matrices of the following dimensions: $\alpha$ is $1 \times K$; $\beta$ is $K \times V$; $\theta$ is $M \times K$; $\phi$ is $K \times V$; $w$ is $M \times N$; and $z$ is $M \times V$.
% To-do: check the dimensions

@@

% Basic LDA
% - Generative process
% - The derivation of the posterior
\par The relationship the between observable and hidden variables, effectively, describes the inference of the underlying topic structure. We will go through an arbitrary example in order to give an intuition in the inference process and familiarise with the lower level notation of the previous listed variables. Figure \ref{fig:lda} shows see that $\alpha_k$ (the prior for the $k$th topic in a document) influences $\theta_{:, k}$ (the latent variable describing the distribution of topic $k$ over all documents). It follows that $\theta_{d, k}$ (the probability of the $k$th topic occurrence in document $d$) influences $z_{d, :}$ (the topic assignment over all vocabulary terms for document $d$). Further, $z_{d, v}$ refers to the topic assignment of vocabulary term $v$. By taking $z_{d, v}$ and $\beta_{v, k}$ (the prior of topic $k$ on vocabulary term $v$), we obtain word $w_{d, v}$. In the complete perspective, the generative process of the corpus is given by the following joint probability distribution
\begin{equation}
p(\beta, \theta, z, w) = \prod_{k=1}^Kp(\beta_{:, k})\prod_{d=1}^Dp(\theta_{d, :})\bigg(\prod_{v=1}^Vp(z_{d, v} | \theta_{d, :})p(w_{d, v} | \beta, z_{d, v})\bigg).
\end{equation}
By using Bayes Theorem, we derive the computation of the conditional distribution
\begin{equation}
p(\beta, \theta, z | w) = \frac{p(\beta, \theta, z, w)}{p(w)}. 
\end{equation}
This equation expresses the computation of the posterior -- the derivation of the hidden corpus structure given observable variable $w$. The problem arises from inefficient analytical computation of evidence $p(w)$. The approximation of it will be considered during the literature review on LDA-like models. The naming of the following subsections corresponds to the reviewed papers.

@@

% Intro
% - Detailed approach in covering implementations details
\par Latent Dirichlet allocation is a probabilistic topic modelling technique proposed by Blei et al. \cite{blei_2003}. Since this paper is the origin of LDA-like models, it will be discussed in more detail to introduce the basis of LDA-like methodology. As a result, this coverage will simplify the review on the LDA derivatives discussed in the upcoming subsections.

@@

% Generative process
% - Three level hierarchy
% - The generative algorithm
\par As discussed in the preliminaries subsection, we are also taking an assumption that the corpus is induced by a probabilistic process. The generation of the corpus has three levels: the parameters $\alpha$ and $\beta$ are sampled once; the variable $\theta$ is sampled once per document; and the variables $z$ and $w$ are sampled once per word. This hierarchical process relates to the generative algorithm given in the original LDA paper, it is equivalent to Algorithm 1 below.
\begin{algorithm}[H]
\caption{Document Generation}
\label{alg:document_generation}
\begin{algorithmic}[1]
\State $N \sim \mbox{Poisson}(\xi)$
\State $\theta \sim \mbox{Dir}(\alpha)$
\For {$n \leftarrow 1, N$}
\State $z_n \sim \mbox{Multinomial}(\theta)$
\State $w_n \sim \mbox{p}(w_n | z_n, \beta)$
\EndFor
\end{algorithmic}
\end{algorithm}
Note that the algorithm describes the generation of a single document. In order to generate the whole corpus, we would run the algorithm for each document. Going back to Algorithm 1, a brief intuition behind it is given as follows: in line 1 we draw the number of words $N$ in a document ($\xi$ is an ancillary variable used as the mean for the Poisson Distribution); in line 2 we draw topic distribution $\theta$ from the Dirichlet distribution on parameter $\alpha$; further, in lines 3--6 we generate the words: in line 4 we draw topic $z_n$ from the multinomial distribution on $\theta$ and, finally, in line 5 we obtain  word $w_n$ from the conditional probability on $z_n$ and $\beta$, recall that $\beta$ is the prior on the vocabulary terms correspondence to topics.

% Variational inference 
% - Variational parameters gamma and fi
\par As suggested in the ending paragraph of the preliminaries subsection, the inference relates to the computation of evidence $p(w)$ (it is the denominator term given in Equation 2). Even though the original paper mentions the inference methods such as Markov chain Monte Carlo, it suggests using the method of variational inference. In order to develop an intuition over the method, the authors introduce the variational parameters $\gamma$ and $\phi$. It is also mentioned that the computation of $\gamma$ and $\phi$ is an optimisation problem. That is, we are assessing the minimum bound over the latter parameters. Further, recall that in Figure \ref{fig:lda} we explicitly express the parameter $\phi$ in describing the basic LDA model. However, the authors have chosen to show it separately -- upon describing the variational method. Their expression is given in Figure \ref{fig:var_par} below.
\begin{figure}[H]
  \centering
  \includegraphics[width=0.4\textwidth]{variational_parameters}
  \caption{Variational parameters}
  \label{fig:var_par}
\end{figure}
Note that the authors provide pseudo code to build an intuition in tackling this optimisation problem.

@@

% Experiments
% - Document modelling: issue with pLSI and mixture unigrams~ 
% - Document classification: 
% -- Dimensionality reduction
% -- Comparison with SVM
\par The conducted experiments display an improvement in performance over the predecessor models and suggest the application domains of LDA-like models. The first experiment addresses document modelling. LDA  has displayed balanced hidden topic proportions in the set of tested documents. That is, the tested documents has replicated the topic proportions of the training documents. The second experiment is directed towards the problem of document classification. They have addressed the \textit{dimensionality reduction} in order to surpass the performance of the support-vector-machine-like models. To be more specific, the documents with reduced dimensionality would possess a lower amount of features. The results of the experiment indicate that the basic LDA model has successfully reduced the number of features. Also, it has displayed an improvement in accuracy compared to the support-vector machine-like model.

% Conclusion
% - Summary of the results established by the paper
\par The review on the original LDA model has set up a basis required to understand the model's derivatives. That is, the introduced methodology will be useful in considering the following improvements: (1) the  relaxation of the exchangeability assumption; (2) the adaptation of time-series suggesting word correspondence to topic over time; (3) the discussion over alternative inference methods. Apart from that, the experiments introduced in this section has reflected the LDA performance upon the model's introduction to the public. The current state-of-the-hard topic modelling techniques overcome the shown results. The experiments were addressed in order to suggest the application fields of LDA-like models.
% To-do: rework the paragraph name to reflect its purpose: remove the discussion (you might think it's done, but it's not)

@@

% Intro
% - Sampling as a method of inference
\par The article titled as `Finding Scientific Topics' \cite{griffiths} suggests an alternative approach to the inference which is suitable for LDA-like models. To be more specific, the article provides a readable introduction on the application of the Markov chain Monte Carlo (MCMC) algorithm. That is, the MCMC algorithm is applied as a sampling-based method to infer an underlying topic structure. Further, the authors discuss the method's application settings for the machine learning problems in vision. Also, the authors conduct an experiment on inferring the structure of topics in the abstracts of papers published in `Proceedings of the National Academy of Sciences' (PNAS). Finally, the paper is concluded by a discussion on assessing the sampling-based method usability. To be more specific, the performance of the sampled-based method is compared to the performance of variational inference methods.

% Gibbs Sampling
% - Recall the meaning of phi and theta.
% - Introduce the inference method
\par The MCMC-based inference method is often referred by \textit{Gibbs sampling}. The authors claim that the method provides a first-order approximation upon inferring the hidden structure of the data. That is, the performance is sufficient to establish quantitative reasoning in assessing the correlating sections and semantic structure of the data. Recall that by $\theta$ we denote the topic distribution over the documents and by $\phi$ we denote the vocabulary terms distribution over the topics. In Gibbs sampling, the latter variables are obtained by examining the posterior distribution, whereas topic assignments $z$ are sampled sequentially. That is, we sample topic assignment $z$ and update the distribution depending on the obtained value (this will impact the next sample of $z$); we repeat this process until the topic distribution converges. To be more specific, each drawn $z_i$ ($i$ denotes the iteration) is recorded as a count. That is, we keep a track of topic assignments for particular words and documents. The value of drawn topic $z_i$ is proportional to the distribution given below:
\begin{equation}
P(z_i = j | z_{-i}, w) \propto \dfrac{n_{-i, j}^{(w_i)} + \beta}{n_{-i, j}^{(\cdot)} + V\beta}\dfrac{n_{-i, j}^{(d_i)} + \alpha}{n_{-i, \cdot}^{(d_i)} + K\alpha}.
\end{equation}
For a brief familiarisation with the notation used in the expression, note the following: $n_j^{d}$ and $n_j^{w}$ are the counts of topic $j$ assignments in document $d$ and word $w$; recall that $V$ is the number of words in the vocabulary and $K$ is the number of topics. Also, note that the counts do not include the last assignment of the topic, this is indicated by $-i$. Ultimately, the left side of the expression is $\phi$ and the right side is $\theta$. That is, 
\begin{align}
\phi_j &= \dfrac{n_{-i, j}^{(w_i)} + \beta}{n_{-i, j}^{(\cdot)} + W\beta},\\
\theta_j &= \dfrac{n_{-i, j}^{(d_i)} + \alpha}{n_{-i, \cdot}^{(d_i)} + K\alpha}.
\end{align}
This process of the topic assignment is executed for every word in the document and for all documents in the corpus. The procedure is repeated until convergence or, more pragmatically, until the distribution displays the assigned topic fluctuations below the set threshold. In the Gibbs sampling procedure, there is no need to explicitly assign $\theta$ and $\phi$ values. If needed, these values are obtained directly from the topic distribution, as shown by the two previous equations.

% Auxilliary experiments
% - Vision
% - Time series
\par The authors has shown a use of Gibbs sampling on a problem of vision. In their experiment settings, it is assumed that a collection of images represents a corpus; an image represent a document; and a pixel of an image is a word. The experiment has shown that the model is able to infer the underlying basis (topics) from which the images could be constructed. The performance of the model was compared to the variational inference based LDA models. Note that the performance was assessed in terms of \textit{perplexity} -- the uncertainty in assigning a word assignment to a topic. It has been shown that Gibbs sampling displays superior performance in the small data sets, whereas the performance in larger data tends to fall off. Furthermore, the authors have assessed the performance upon varying the number of topics. The results of this experiment indicate a trade-off in deriving high-level and low-level structures. That is, the low number of topics might result in mashing several structures, whereas the high number of topics is likely to infer irrelevant structures.

% The main experiment
% - PNAS
\par The main experiment of the article is conducted on the PNAS data set containing abstracts of the articles published from 1991 to 2001. The authors have noticed that the topic distribution over the vocabulary terms change over time. However, the model had not yet been optimised to take this notion of time-series into account upon the inference. Nevertheless, the authors have emphasised the necessity to assess the dynamic topic modelling.

@@

% Intro:
% - Time series
% - Qualitative and quantitative improvement
\par At this point, we have set up the preliminaries to cover the article on utilising the notion of time-series in LDA-like models. We review the initial article on utilising time-series -- `Dynamic Topic Models' \cite{blei_2006}. Note that by a \textit{static} model we mean that the assumed generative process does not take the development/change of the vocabulary terms into account, whereas in \textit{dynamic} topic modelling we assume the notion of that the documents were generated in a sequential process. That is, in dynamic topic modelling, the topics evolve in terms of their distribution over the vocabulary terms. Upon reviewing the dynamic topic modelling article, we emphasise the following aspects: the fundamental dynamic topic modelling assumptions and their differences compared to the basic LDA model; the generative process; the suggested inference methods; and the results displayed by the experiments on the initial dynamic topic model.
% To-do: qualitative and quantitative improvements

% Assumptions
% - Data is divided in time slices
\par The assumptions set for the dynamic topic model emphasises its strengths over the basic LDA model. Most importantly, the dynamic topic model relaxes the assumption of the document exchangeability. Since the documents are generated over time, we assign them to specific time-slices. The combination of these time-slices would represent the corpus. Note that the documents in each time-slice are exchangeable. This treatment of time set ups a basis to tackle the problems which involve the sequential development of the data. On the other hand, note that another assumption of discrete data remains the same as in the basic LDA model. Even though we are inducing time series (which would suggest the use of \textit{continuous} variables), the authors use only \textit{categorical} (discrete) data. The distinction between continuous and categorical data is the following: continuous data is strictly numeric and infinite, whereas categorical data takes a fixed value from a finite set of values.

@@

% Section intro
% - Discussion over literature
% - Design of the model
% - Evaluation of the model
\par In this section, we will set up the approach of the project's execution. To start with, we will go over the literature reviewed in Section 3 to derive the promising techniques in terms of applicability to MSI data sets. Then, we will go over the high-level design and implementation requirements. Finally, we will set the basis for the model's evaluation.

@@

\begin{figure}[H]
\begin{ganttchart}[
hgrid, 
vgrid,
y unit chart=.8cm
]{1}{15}

@@

\end{ganttchart}
\caption{Gantt chart of the project's execution.}
\label{fig:gantt}
\end{figure}

>>>>

% Intro
% - ML in biology
% - MSI
% - Spatial Smoothing
\par This proposal is focused on applying machine learning techniques in order to improve the analysis of biology data sets. We intend to enhance the quality of \textit{mass spectrometry imaging} (MSI) data. Briefly, by the use of MSI, the researchers are able to interpret the formation of chemical processes. However, as suggested by a recent overview by Smith et al. \cite{smith_2004}, the pre-processing of molecular-level entity data sets is lagging in terms of the current noise reduction algorithm effectiveness. We will attempt to tackle this issue of noise by utilising spatial smoothing.

@@

% Objectives
% - Research in the state-of-the-art methodology.
% - Reducing the impact of noise in metabolomics.
% - Providing a general-purpose machine learning model.
\par The primary goal of this research is to expand the consensus on whether topic modelling is a valid approach to apply spatial smoothing in MSI data sets. To be more specific, we set our objectives to the following:
\begin{enumerate}
    \item Optimise the general topic modelling methods for MSI data sets;
    \item Successfully apply spatial smoothing in MSI data sets;
    \item Give a basis to a general topic modelling method for error correction.
\end{enumerate}
Briefly, for the first objective, we will conduct a survey on LDA-like topic modelling methodology. This survey will distinguish the potential models for the applications in MSI. The second objective focuses on optimising these potential models for the particular task of spatial smoothing in MSI data sets. Finally, the third objective can be treated as a general contribution to the field of machine learning, since the discovered spatial smoothing techniques might be applicable in other domains.

@@

% General intuition on topic modelling
% - Example
\par Recall that topic modelling infers hidden patterns from the processed data sets. For example, if we were analysing a book, its chapters might address different topics or address the topics at different extent. Further, the chapters itself would have unique distributions of the topics. Note that we are assuming that there is no knowledge about the actual topics covered in the book. In order to derive these hidden topics, a probabilistic topic modelling method would analyse the structure of the words used in the book. In other words, the model would induce patterns which would refer to a particular topic. Regarding the method's performance, it would depend on the method's ability to optimise the inference of the patterns. For example, commonly used words (such as and, or, the, etc.) would be relevant in all topics. Therefore, their semantic impact in defining the topics would be negligible. For this reason, we could dismiss these to improve the time performance of the method.

% LDA intro
% - the model
% - A notification that the terminology and the notation will be useful throughout the proposal.
\par The terminology used in this proposal is similar to the terminology used by Blei \cite{blei_2012}. Figure \ref{fig:lda} below represents the original LDA model in plate notation. We will use it as a guide to familiarise with the terminology.   
\begin{figure}[H]
  \centering
  \includegraphics[width=0.7\textwidth]{lda_model}
  \caption{The design of the original LDA model}
  \label{fig:lda}
\end{figure}

% High level LDA terminology
% - Notation:
% -- List LDA definitions
% -- Explain the definitions
\par To start with, we will familiarise with the higher level entities of the model. The circles represent an entity, whereas the plates indicate the number of entities. For example, there would be $N \times M$ entities of $z$ and $M$ entities of $\theta$. Further, LDA is a three-level hierarchical structure: the smallest entity (grey circle) is defined as a \textit{word}; then a collection of words (inner plate) is defined as a \textit{document}; and a collection of documents (outer plate) is a \textit{corpus}. Note that Figure \ref{fig:lda} represents $M$ documents and $N$ words per document, it follows that in the corpus there would be $M \times N$ words. Also, as suggested by Figure \ref{fig:lda}, let's denote a word by $w$; then, the sequence of words in document $d$ is denoted by 
\begin{equation*}
\mathbf{w}_d = \{w_1, w_2, \dots, w_N\},
\end{equation*}
and the sequence of documents in the corpus is denoted by
\begin{equation*}
\mathcal{\mathbf{D}} = \{\mathbf{w_1}, \mathbf{w_2}, \dots, \mathbf{w_M}\}.
\end{equation*}

% Low level LDA terminology
% - Intro to observable and hidden variables
% - Notation:
% -- List the definitions
% -- Explain the definitions
\par Now, we will introduce the lower level entities. The grey circle indicates the \textit{observable} variables, whereas the white coloured circles indicate the \textit{hidden} variables. Note that in Figure 1 the words are the only observable variables; all other entities are hidden, i.e., they describe the underlying structure of the model. Further, recall that topic modelling is applied in order to discover a mixture of topics. In other words, topic modelling could describe a corpus and its documents in terms of topic distributions. This intuition will simplify the definitions of the remaining entities in Figure 1, these are given as follows:
\begin{itemize}
\item $K$ is the number of topics;
\item $V$ is the size of the vocabulary, i.e. the number of unique words in the corpus;
\item $\alpha$ is the parameter referring to the prior distribution of the topics over documents;
\item $\beta$ is the parameter referring to the prior distribution of the vocabulary over topics;
\item $\theta$ is the latent variable referring to the topic distribution over documents;
\item $\phi$ is the latent variable referring to the vocabulary over topics
\item $w$ is the observable variable referring to the words in documents;
\item $z$ is the latent variable of topic assignments over the words.
\end{itemize}
Further, note that the listed entities can be perceived as matrices of the following dimensions: $\alpha$ is $1 \times K$; $\beta$ is $K \times V$; $\theta$ is $M \times K$; $\phi$ is $K \times V$; $w$ is $M \times N$; and $z$ is $M \times V$.
% To-do: check the dimensions

@@

% Basic LDA
% - Generative process
% - The derivation of the posterior
\par The relationship of observable and hidden variables, effectively, describes the inference of the underlying topic structure. We will go through an arbitrary example in order to give an intuition in the inference process; also, we will familiarise with the lower level notation of the previous listed variables. Figure \ref{fig:lda} illustrates that $\alpha_k$ (the prior for the $k$th topic in a document) influences $\theta_{:, k}$ (the latent variable describing the distribution of topic $k$ over all documents). It follows that $\theta_{d, k}$ (the probability of the $k$th topic occurrence in document $d$) influences $z_{d, :}$ (the topic assignment over all vocabulary terms for document $d$). Further, $z_{d, v}$ refers to the topic assignment of vocabulary term $v$. By taking $z_{d, v}$ and $\beta_{v, k}$ (the prior of topic $k$ on vocabulary term $v$), we obtain word $w_{d, v}$. In the complete perspective, the generative process of the corpus is given by the following joint probability distribution
\begin{equation}
p(\beta, \theta, z, w) = \prod_{k=1}^Kp(\beta_{:, k})\prod_{d=1}^Dp(\theta_{d, :})\bigg(\prod_{v=1}^Vp(z_{d, v} | \theta_{d, :})p(w_{d, v} | \beta, z_{d, v})\bigg).
\end{equation}
By using Bayes Theorem, we derive the computation of the conditional distribution
\begin{equation}
p(\beta, \theta, z | w) = \frac{p(\beta, \theta, z, w)}{p(w)}. 
\end{equation}
This equation expresses the computation of the posterior -- the derivation of the hidden corpus structure given observable variable $w$. The problem arises from inefficient analytical computation of evidence $p(w)$. The approximation of it will be considered during the literature review on LDA-like models. Note that the naming of the following subsections corresponds to the reviewed papers.

@@

% Intro
% - Detailed approach in covering implementations details
\par Latent Dirichlet allocation is a probabilistic topic modelling technique proposed by Blei et al. \cite{blei_2003}. Since this paper is the origin of LDA-like models, it will be discussed in more detail in order to introduce the basis of LDA-like methodology. As a result, this coverage will simplify the review on the LDA derivatives discussed in the upcoming subsections.

@@

% Generative process
% - Three level hierarchy
% - The generative algorithm
\par As discussed in the preliminaries subsection, we also set an assumption that the corpus is induced by a probabilistic process. The generation of the corpus has three levels: the parameters $\alpha$ and $\beta$ are sampled once; the variable $\theta$ is sampled once per document; and the variables $z$ and $w$ are sampled once per word. This hierarchical process relates to the generative algorithm given in the original LDA paper, it is equivalent to Algorithm 1 below.
\begin{algorithm}[H]
\caption{Document Generation}
\label{alg:document_generation}
\begin{algorithmic}[1]
\State $N \sim \mbox{Poisson}(\xi)$
\State $\theta \sim \mbox{Dir}(\alpha)$
\For {$n \leftarrow 1, N$}
\State $z_n \sim \mbox{Multinomial}(\theta)$
\State $w_n \sim \mbox{p}(w_n | z_n, \beta)$
\EndFor
\end{algorithmic}
\end{algorithm}
Note that the algorithm describes the generation of a single document. In order to generate the whole corpus, we would run the algorithm for each document. Going back to Algorithm 1, a brief intuition behind it is given as follows: in line 1 we draw the number of words in a document $N$($\xi$ is an ancillary variable used as the mean for the Poisson Distribution); in line 2 we draw topic distribution $\theta$ from the Dirichlet distribution on parameter $\alpha$; further, in lines 3--6 we generate the words: in line 4 we draw topic $z_n$ from the multinomial distribution on $\theta$ and, finally, in line 5 we obtain  word $w_n$ from the conditional probability on $z_n$ and $\beta$, recall that $\beta$ is the prior on the vocabulary terms correspondence to topics.

% Variational inference 
% - Variational parameters gamma and fi
\par As suggested by the last paragraph of the preliminaries subsection, the inference relates to the computation of evidence $p(w)$ (it is the denominator term given in Equation 2). Even though the original paper mentions the inference methods such as Markov chain Monte Carlo, the authors suggest using the method of variational inference. In order to develop an intuition over the method, the authors introduce the variational parameters $\gamma$ and $\phi$. It is also mentioned that the computation of $\gamma$ and $\phi$ is an optimisation problem. That is, we are assessing the minimum bound over the latter parameters. Further, recall that in Figure \ref{fig:lda} we explicitly express the parameter $\phi$ in describing the basic LDA model. However, the authors have chosen to show it separately -- upon describing the variational method. Their expression is given in Figure \ref{fig:var_par} below.
\begin{figure}[H]
  \centering
  \includegraphics[width=0.4\textwidth]{variational_parameters}
  \caption{Variational parameters}
  \label{fig:var_par}
\end{figure}
Note that the authors provide pseudo code to build an intuition in tackling this optimisation problem.

@@

% Experiments
% - Document modelling: issue with pLSI and mixture unigrams~ 
% - Document classification: 
% -- Dimensionality reduction
% -- Comparison with SVM
\par The conducted experiments display an improvement in performance over the predecessor models and suggest the application domains of LDA-like models. The first experiment addresses document modelling. To be more specific, LDA has displayed balanced hidden topic proportions in the set of tested documents. That is, the tested documents has replicated the topic proportions of the training documents. The second experiment is directed towards the problem of document classification. The authors have addressed the \textit{dimensionality reduction} in order to surpass the performance of the support vector machine (SVM) models. To be more specific, the documents with reduced dimensionality would possess a lower amount of features. The results of the experiment indicate that the basic LDA model has successfully reduced the number of features. Also, it has displayed an improvement in accuracy compared to the SVM-like models.

% Conclusion
% - Summary of the results established by the paper
\par The review on the original LDA model has set up a basis required to understand the model's derivatives. That is, the introduced methodology will be useful in considering the following improvements: (1) the  relaxation of the exchangeability assumption; (2) the adaptation of time-series suggesting the word-topic correspondence over time; (3) the discussion over alternative inference methods. Apart from that, the experiments introduced in this section have reflected the LDA performance during the time of the initial model release. The current state-of-the-hard topic modelling techniques overcome the shown results. Nevertheless, the carried experiments suggest the application fields of LDA-like models.
% To-do: rework the paragraph name to reflect its purpose: remove the discussion (you might think it's done, but it's not)

@@

% Intro
% - Sampling as a method of inference
\par The research paper titled as `Finding Scientific Topics' \cite{griffiths} suggests an alternative approach to the inference which is suitable for LDA-like models. To be more specific, the article provides a readable introduction on the application of the Markov chain Monte Carlo (MCMC) algorithm. That is, the MCMC algorithm is applied as a sampling-based method to infer an underlying topic structure. Further, the authors discuss the method's application settings for the machine learning problems in vision. Also, the authors conduct an experiment on inferring the structure of topics in the abstracts of papers published in `Proceedings of the National Academy of Sciences' (PNAS). Finally, the paper is concluded by a discussion on assessing the sampling-based method usability. To be more specific, the performance of the sampled-based method is compared to the performance of variational inference methods.

% Gibbs Sampling
% - Recall the meaning of phi and theta.
% - Introduce the inference method
\par The MCMC-based inference method is often referred by \textit{Gibbs sampling}. The authors claim that the method provides a first-order approximation for the inference of the hidden structure of the data. That is, the performance is sufficient to establish quantitative reasoning in assessing the correlating sections and semantic structure of the data. Recall that by $\theta$ we denote the topic distribution over the documents and by $\phi$ we denote the vocabulary terms distribution over the topics. In Gibbs sampling, the latter variables are obtained by examining the posterior distribution, whereas topic assignments $z$ are sampled sequentially. That is, we sample topic assignment $z$ and update the distribution depending on the obtained value (this will impact the next sample of $z$); we repeat this process until the topic distribution converges. To be more specific, each drawn $z_i$ ($i$ denotes the iteration) is recorded as a count. That is, we keep a track of topic assignments for particular words and documents. The value of drawn topic $z_i$ is proportional to the distribution given below:
\begin{equation}
P(z_i = j | z_{-i}, w) \propto \dfrac{n_{-i, j}^{(w_i)} + \beta}{n_{-i, j}^{(\cdot)} + V\beta}\dfrac{n_{-i, j}^{(d_i)} + \alpha}{n_{-i, \cdot}^{(d_i)} + K\alpha}.
\end{equation}
For a brief familiarisation with the notation used in the expression, note the following: $n_j^{d}$ and $n_j^{w}$ are the counts of topic $j$ assignments in document $d$ and word $w$; recall that $V$ is the number of words in the vocabulary and $K$ is the number of topics. Also, note that the counts do not include the last assignment of the topic, this is indicated by $-i$. Ultimately, the left side of the expression is $\phi$ and the right side is $\theta$. That is, 
\begin{align}
\phi_j &= \dfrac{n_{-i, j}^{(w_i)} + \beta}{n_{-i, j}^{(\cdot)} + W\beta},\\
\theta_j &= \dfrac{n_{-i, j}^{(d_i)} + \alpha}{n_{-i, \cdot}^{(d_i)} + K\alpha}.
\end{align}
This process of the topic assignment is executed for every word in the document and for all documents in the corpus. The procedure is repeated until convergence or, more pragmatically, until the distribution displays the assigned topic fluctuations below the set threshold. In the Gibbs sampling procedure, there is no need to explicitly assign $\theta$ and $\phi$ values. If needed, these values are obtained directly from the topic distribution, as shown by the two previous equations.

% Auxilliary experiments
% - Vision
% - Time series
\par The authors have shown an application of Gibbs sampling in a problem of vision. In their experiment settings, it is assumed that a collection of images represents a corpus; an image represents a document; and a pixel of an image is a word. The experiment has shown that the model is able to infer the underlying basis (topics) from which the images could be constructed. The performance of the model was compared to the variational inference based LDA models. Note that the performance was assessed in terms of \textit{perplexity} -- the uncertainty in assigning a word assignment to a topic. It has also been shown that Gibbs sampling displays superior performance in the small data sets, whereas the performance in larger data sets tends to fall off. Furthermore, the authors have assessed the performance upon varying the number of topics. The results of this experiment indicate a trade-off in deriving high-level and low-level patterns. That is, the low number of topics might result in mashing significant patterns, whereas the high number of topics is likely to infer irrelevant patterns.

% The main experiment
% - PNAS
\par The main experiment of the research paper is conducted on the PNAS data set containing abstracts of the articles published from 1991 to 2001. The authors have noticed that the topic distribution over the vocabulary terms change over time. However, the model had not yet been optimised to take this notion of time-series into account upon the inference. Nevertheless, the authors have emphasised the necessity to assess the dynamic topic modelling.

@@

% Intro:
% - Time series
% - Qualitative and quantitative improvement
\par At this point, we have set up the preliminaries to cover the research paper on utilising the notion of time-series in LDA-like models. That is, we review the initial paper on utilising time-series in LDA-like models -- `Dynamic Topic Models' \cite{blei_2006}. Note that by a \textit{static} model we mean that the assumed generative process does not take the development/change of the vocabulary terms into account, whereas in \textit{dynamic} topic modelling we assume the notion that the documents were generated in a sequential process. That is, in dynamic topic modelling, the topics evolve in terms of their distribution over the vocabulary terms. By reviewing this research paper, we emphasise the following aspects: the fundamental dynamic topic modelling assumptions and their differences compared to the basic LDA model; the generative process; the suggested inference methods; and the results displayed by the experiments on the initial dynamic topic model.
% To-do: qualitative and quantitative improvements

% Assumptions
% - Data is divided in time slices
\par The assumptions set for the dynamic topic model emphasises its strengths over the basic LDA model. Most importantly, the dynamic topic model relaxes the assumption of the document exchangeability. Since the documents are generated over time, we assign them to specific time-slices. The combination of these time-slices would represent the corpus. Note that the documents in each time-slice are exchangeable. This treatment of time sets a basis to tackle the problems which involve the sequential development of the data. On the other hand, note that another assumption of discrete data remains the same as in the basic LDA model. Even though we are inducing time series (which would suggest the use of \textit{continuous} variables), the authors use only \textit{categorical} (discrete) data. The distinction between continuous and categorical data is the following: continuous data is strictly numeric and infinite, whereas categorical data takes a fixed value from a finite set of values.

@@

% Section intro
% - Discussion over literature
% - Design of the model
% - Evaluation of the model
\par In this section, we will set up the approach of the project's execution. To start with, we will go over the literature reviewed in Section 3 to derive the promising techniques in terms of applicability to MSI data sets. Then, we will go over the high-level design and implementation requirements. Finally, we will set the basis for the model's evaluation and assess the potential risks.

@@

\begin{figure}[H]
\begin{ganttchart}[
hgrid, 
vgrid,
y unit chart=.7cm
]{1}{15}

@@

\end{ganttchart}
\caption{The Gantt chart estimating the project's execution.}
\label{fig:gantt}
\end{figure}
